---
title: "Lab 03 - Parametric and Nonparametric Univariate Tests"
date: 2021/10/8
author: "Nonparametric statistics ay 2021/2022"
output:
  
  html_document: 
    df_print: paged
  pdf_document: default
  html_notebook: 
    df_print: paged
  word_document: default
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{css, echo=FALSE}
.extracode {
background-color: lightblue;
}
```

# Nonparametric Statistics and Functional Data
The Objective of this brief lab is to show you some applications of the advanced techniques you have seen during the course, applied to functional data. Namely we will see permutation testing (both global and local...), and prediction

Let's load the packages we need and our data:

```{r}
library(fda)
data=growth #data from the berkeley growth study...
```

And let's plot my curves...

```{r}
matplot(data$age,data$hgtm, type='l',col='blue')
matlines(data$age,data$hgtf, type='l',col='red')
```

##Permutation Global Testing
What if I want to test if the two curves are equal or not? Nothing simpler.. I just need to remember how permutation tests work...

```{r}
seed=2781991
B=1000
berkeley=rbind(t(data$hgtm),t(data$hgtf))
n=nrow(berkeley)
n_m=nrow(t(data$hgtm))
n_f=nrow(t(data$hgtf))


meandiff=(colMeans(t(data$hgtm))-colMeans(t(data$hgtf)))
plot(meandiff,type = 'l')
T0=sum(meandiff^2)
T0
```
And, Knowing that under $H_0$ the two groups of curves are IID, my likelihood-invariant permutation scheme is of course label permutation, so...

```{r}
T0_perm=numeric(B)

for(perm in 1:B){
  permutazione <- sample(n)
  berkeley_perm=berkeley[permutazione,]
  perm_m = berkeley_perm[1:n_m,] 
  perm_f = berkeley_perm[(n_m+1):n,] 
  T0_perm[perm]=sum(((colMeans(perm_m)-colMeans(perm_f)))^2)
}

sum(T0_perm >= T0)/B
hist(T0_perm,xlim = c(0,2000))
abline(v=T0,col='green')

```

What would have happened instead, if I were to test inside a group?

```{r}
male1=berkeley[1:(n_m/2),]
male2=berkeley[(n_m/2):n_m,]
ber_m=rbind(male1,male2)

T0=sum(((colMeans(male1)-colMeans(male2)))^2)
T0

T0_perm=numeric(B)

for(perm in 1:B){
  permutazione <- sample(n_m)
  berkeley_perm=ber_m[permutazione,]
  perm_m = berkeley_perm[1:(n_m/2),] 
  perm_f = berkeley_perm[(n_m/2):n_m,] 
  T0_perm[perm]=sum(((colMeans(perm_m)-colMeans(perm_f)))^2)
}

sum(T0_perm >= T0)/B
hist(T0_perm)
abline(v=T0,col='green')
```
Expectedly, I am not rejecting the null hypothesis (Pvalue of the test is very high...)

Now, what I am doing here is basically is testing the hypothesis globally, I am rejecting if, for at least one time instant $t$ the two curves are statistically different.
How do I tell what is that specific time instant? I use a procedure called Inteval-wise Testing

Let's load the package

```{r}
#devtools::install_github("alessiapini/fdatest")
library(fdatest)
```

And let's run the test: I will show just a 2 sample case, very simple and straightforward...

```{r}

tst=IWT2(t(data$hgtm),t(data$hgtf))
plot(tst)

```


In this way, I can see that I can reject the null hypothesis of the two curves being equal, THANKS to the contribution of the latest years...

How do I instead, PREDICT functional data? You've guessed it right, I can use Conformal Prediction

let's try to generate a prediction band for a new male...
I need of course to work on a split conformal setting, so.

```{r}
alpha=.1
n=nrow(ber_m)
i1=sample(1:n,n/2)
t_set=ber_m[i1,]
c_set=ber_m[-i1,]
mu=colMeans(t_set)
res=c_set-mu
ncm=apply(res,2,max)
ncm_sort=c(sort(ncm),Inf)
d=ncm_sort[ceiling((n/2 + 1)*(1-alpha))]

matplot(cbind(mu,mu+d,mu-d),type='l')

```

What if I want to scale the amplitude of this?

```{r}


S=apply(t_set,2,var)
res=(c_set-mu)/S
ncm=apply(res,2,max)
ncm_sort=c(sort(ncm),Inf)
d=ncm_sort[ceiling((n/2 + 1)*(1-alpha))]

matplot(cbind(mu,mu+(d*S),mu-(d*S)),type='l')
```

